Exploring dimensions of biodiversity in Japanese ferns

### 2021-08-26 ms/SI.Rmd
Don't include raw occurrence data in _targets.R
Major refactor. Run process_raw_data.R  first to output cleaned data to data/, then run _targets.R

```{r load-data-si, cache = FALSE}
tar_load(c(
  "japan_shp",
  "comm_scaled_list",
  "occ_point_data_ferns",
  "biodiv_ferns_spatial",
  "biodiv_ferns_endemic_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "traits_nmds",
  "ppgi",
  "phy_sig_results",
  "binary_sig_results",
  "lrt_results",
  "japan_shp",
  "morans_i",
  "t_test_results",
  "aic_env_repro",
  "comm_ferns",
  "repro_data",
  "japan_pterido_tree",
  "model_fits_not_ult",
  "biodiv_ferns_cent"
))
```

@@

```{r map-plot-prep}
# Download high-res map of Japan,
# crop to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))

@@

```{r sampling-curve, fig.width = 4, fig.height = 3}
# Format fern abundances (no hybrids) as numeric vector
fern_abun <-
  occ_point_data_ferns %>%
  group_by(taxon) %>%
  count(sort = TRUE) %>%
  pull(n)

# Run iNEXT
inext_out <- iNEXT::iNEXT(fern_abun, q=0, datatype="abundance")

# create subset of observed points to add to plot as dots
observed_data <- inext_out$iNextEst[inext_out$iNextEst$method == "observed",]

# check estimated completeness
est_completeness <- inext_out$iNextEst %>% filter(method == "observed") %>% pull(SC) %>% scales::percent()
rich_interval <- glue::glue("{observed_data$qD.LCL %>% round} to {observed_data$qD.UCL %>% round}")
completeness_interval <- glue::glue("{observed_data$SC.LCL %>% scales::percent()} to {observed_data$SC.UCL %>% scales::percent()}")

# Make iNEXT plot
inext_out$iNextEst %>%
  filter(method != "observed") %>%
  ggplot(aes(x=m, y=qD)) +
  geom_ribbon(aes(ymin = qD.LCL, ymax = qD.UCL), fill = "grey70") +
  geom_line(aes(linetype=method), size=0.8) +
  scale_linetype_manual(values=c("dotted", "solid", "solid")) +
  geom_point(data=observed_data, size=3) +
  scale_y_continuous("Richness") +
  scale_x_continuous("No. specimens") +
  theme_minimal() +
  theme(
    legend.position="bottom", 
    legend.title=element_blank()
  )
```

@@

# Combine results of building community matrices at different grain sizes
redundancy_by_res <- 
  list(
    "0.1" = comm_scaled_list %>% filter(resol == 0.1) %>% pull(poly_shp) %>% pluck(1),
    "0.2" = comm_scaled_list %>% filter(resol == 0.2) %>% pull(poly_shp) %>% pluck(1),
    "0.3" = comm_scaled_list %>% filter(resol == 0.3) %>% pull(poly_shp) %>% pluck(1),
    "0.4" = comm_scaled_list %>% filter(resol == 0.4) %>% pull(poly_shp) %>% pluck(1)
  ) %>%
  map_df(~as_tibble(.) %>% select(grids, abundance, richness), .id = "res") %>%
  # Calculate redundancy
  mutate(redundancy = 1 - (richness/abundance))

# Summarize redundancy statistics by grain size
redundancy_summary <-
  redundancy_by_res %>%
  mutate(above_50 = ifelse(redundancy > 0.5, 1, 0)) %>%
  mutate(singleton = ifelse(redundancy == 0, 1, 0)) %>%
  group_by(res) %>%
  summarize(
    n_total = n(),
    n_above_50 = sum(above_50),
    n_singletons = sum(singleton),
    percent_above_50 = n_above_50/n_total,
    percent_singletons = n_singletons/ n_total,
    .groups = "drop"
  )

>>>>

```{r load-data-si, cache = FALSE}
tar_load(c(
  "japan_shp",
  "inext_res",
  "redundancy_by_res",
  "redundancy_summary",
  "biodiv_ferns_spatial",
  "biodiv_ferns_endemic_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "traits_nmds",
  "ppgi",
  "phy_sig_results",
  "binary_sig_results",
  "lrt_results",
  "japan_shp",
  "morans_i",
  "t_test_results",
  "aic_env_repro",
  "comm_ferns",
  "repro_data",
  "japan_pterido_tree",
  "model_fits_not_ult",
  "biodiv_ferns_cent"
))
```

@@

```{r map-plot-prep}
# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))

@@

```{r sampling-curve, fig.width = 4, fig.height = 3}
# create subset of observed points to add to plot as dots
observed_data <- inext_res[inext_res$method == "observed",]

# check estimated completeness
est_completeness <- inext_res %>% filter(method == "observed") %>% pull(SC) %>% scales::percent()
rich_interval <- glue::glue("{observed_data$qD.LCL %>% round} to {observed_data$qD.UCL %>% round}")
completeness_interval <- glue::glue("{observed_data$SC.LCL %>% scales::percent()} to {observed_data$SC.UCL %>% scales::percent()}")

# Make iNEXT plot
inext_res %>%
  filter(method != "observed") %>%
  ggplot(aes(x=m, y=qD)) +
  geom_ribbon(aes(ymin = qD.LCL, ymax = qD.UCL), fill = "grey70") +
  geom_line(aes(linetype=method), size=0.8) +
  scale_linetype_manual(values=c("dotted", "solid", "solid")) +
  geom_point(data=observed_data, size=3) +
  scale_y_continuous("Richness") +
  scale_x_continuous("No. specimens") +
  theme_minimal() +
  theme(
    legend.position="bottom", 
    legend.title=element_blank()
  )
```

@@

# Summarize redundancy statistics by grain size
redundancy_summary <- sum_redundancy_by_res(redundancy_by_res)

### 2021-08-26 ms/manuscript.Rmd
Don't include raw occurrence data in _targets.R
Major refactor. Run process_raw_data.R  first to output cleaned data to data/, then run _targets.R

```{r load-data}
# Load objects from targets workflow
tar_load(c(
  "repro_data",
  "occ_point_data_ferns",
  "occ_point_data_ferns_unfiltered",
  "biodiv_ferns_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "signif_cells_protected_area",
  "protected_areas",
  "japan_shp",
  "phy_sig_results",
  "binary_sig_results",
  "traits_summary",
  "fern_traits",
  "japan_pterido_tree",
  "japan_fern_tree",
  "ppgi",
  "occ_point_data_ferns",
  "comm_ferns_full",
  "comm_ferns",
  "japan_points_raw",
  "plastome_calibration_dates",
  "plastome_tree",
  "green_list",
  "model_params",
  "lrt_results",
  "aic_env_repro",
  "model_fits",
  "biodiv_ferns_cent",
  "refs_yaml",
  "refs_other_yaml",
  "template_file",
  "csl_file",
  "broad_alignment_list",
  "japan_rbcL",
  "ms_functions"
))
```

@@

FIXME: add acknowledgements (not including reviewers) JSPS KAKENHI Grant Number 16H06279

@@

@@

**Methods** We used data from >`r nrow(occ_point_data_ferns) %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa. 
We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode. 
We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism. 
We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. 
We compared the overlap of significantly diverse grid-cells with existing conservation zones to assess conservation status.

@@

We used a list of all non-hybrid fern specimens housed at herbaria in Japan. 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated). 
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html).

@@

```{r dataset-stats}
### Taxon list ###
# Make list of all native ferns in Japan (based off Green List), includes hybrids
fern_list <-
  green_list %>%
  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
  left_join(ppgi) %>%
  filter(class == "Polypodiopsida")

### Phylogeny ###
total_fern_taxa <- fern_list %>%
  filter(hybrid == FALSE) %>%
  pull(taxon) %>%
  n_distinct()

phylogeny_sampled_percent <- ape::Ntip(japan_fern_tree) %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)

@@

### Specimens ###
n_specimens <- nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")
n_specimens_taxa <- occ_point_data_ferns %>% pull(taxon) %>% n_distinct()
n_specimens_taxa_percent <- n_specimens_taxa %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)

@@

```{r distribution-stats}
# Calculate number of genera and families restricted to lat < 30.1
southern_genera <-
  occ_point_data_ferns %>%
  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
  group_by(genus) %>%
  summarize(
    max_lat = max(latitude),
    .groups = "drop"
  ) %>%
  filter(max_lat < 30.1)

n_southern_genera <- southern_genera %>% pull(genus) %>% n_distinct %>% english2

southern_families <-
  occ_point_data_ferns %>%
  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
  left_join(ppgi, by = "genus") %>%
  group_by(family) %>%
  summarize(
    max_lat = max(latitude),
    .groups = "drop"
  ) %>%
  filter(max_lat < 30.1)

n_southern_families <- southern_families %>% pull(family) %>% n_distinct %>% english2
```

@@

```{r limits-stats}
# Summarize occurrences by latitude
occ_data_by_lat <-
  occ_point_data_ferns %>%
  group_by(taxon) %>%
  summarize(
    min_lat = min(latitude),
    max_lat = max(latitude)
  ) %>%
  mutate(
    lat_span = max_lat - min_lat
  )

# Find the species with the widest latitudinal distribution
taxon_with_widest_lat <- occ_data_by_lat %>%
  slice_max(lat_span) %>%
  left_join(green_list, by = "taxon") %>%
  taxastand::add_parsed_names(sci_name = scientific_name, species, author) %>%
  mutate(text = glue("*{species}* {author}"))

# Find number of species occurring across areas defined by straits
n_widespread_taxa <-
  occ_data_by_lat %>%
  # Filter by approx location of Tokara strait (29) and Tsugaru Strait (41.5)
  filter(min_lat < 29, max_lat > 41.5) %>%
  pull(taxon) %>%
  n_distinct %>%
  english2

>>>>

```{r load-data}
# Load objects from targets workflow
tar_load(c(
  "aic_env_repro",
  "binary_sig_results",
  "biodiv_ferns_cent",
  "biodiv_ferns_spatial",
  "broad_alignment_list",
  "comm_ferns",
  "comm_ferns_full",
  "csl_file",
  "fern_traits",
  "green_list",
  "japan_fern_tree",
  "japan_points_raw",
  "japan_pterido_tree",
  "japan_rbcL",
  "japan_shp",
  "k_phylogeny",
  "k_taxonomy",
  "lat_span_summary",
  "lrt_results",
  "model_fits",
  "model_params",
  "ms_functions",
  "occ_point_data_summary",
  "phy_sig_results",
  "plastome_calibration_dates",
  "plastome_tree",
  "ppgi",
  "protected_areas",
  "refs_other_yaml",
  "refs_yaml",
  "repro_data",
  "signif_cells_protected_area",
  "template_file",
  "traits_summary"
))
```

@@

Members of the Iwasaki Lab at the University of Tokyo provided comments that improved the manuscript.
Shawn Laffan provided advice about CANAPE.
This study was supported by JSPS KAKENHI Grant Number 16H06279.

@@

```{r abstract-stats}
# Convert occ_point_data_summary dataframe to nested list for easier indexing
occ_sum <- occ_point_data_summary %>%
  # change dataset name to just "unfiltered" vs. "filtered"
  mutate(dataset = case_when(
    str_detect(dataset, "unfiltered") ~ "unfiltered",
    TRUE ~ "filtered")) %>%
  split(.$dataset) %>%
  map(., ~split(., .$variable))
```

@@

**Methods** We used data from >`r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa. 
We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode. 
We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism. 
We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. 
We compared the overlap of significantly diverse grid-cells with existing conservation zones to assess conservation status.

@@

We used a list of all non-hybrid fern specimens housed at herbaria in Japan. 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated). 
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).

@@

```{r dataset-stats}
### Taxon list ###
# Make list of all native ferns in Japan (based off Green List), *including* hybrids
fern_list <-
  green_list %>%
  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
  left_join(ppgi) %>%
  filter(class == "Polypodiopsida")

# Caculate number of native, non-hybrid fern taxa
total_fern_taxa <- fern_list %>%
  filter(hybrid == FALSE) %>%
  pull(taxon) %>%
  n_distinct()

### Phylogeny ###
phylogeny_sampled_percent <- ape::Ntip(japan_fern_tree) %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)

@@

### Specimens (filtered occurrence data) ###
n_specimens <- occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")
n_specimens_taxa <- occ_sum$filtered$n_taxa$value
n_specimens_taxa_percent <- n_specimens_taxa %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)

@@

```{r distribution-stats}
# Summarize maximum latitude by genus, family
max_lat_summary <- summarize_fern_lat_max(lat_span_summary, ppgi)

# Calculate number of genera and families restricted to lat < 30.1
southern_genera <-
  max_lat_summary %>%
  filter(max_lat < 30.1, taxon_level == "genus")

n_southern_genera <- southern_genera %>% pull(higher_taxon) %>% n_distinct %>% english2

southern_families <-
  max_lat_summary %>%
  filter(max_lat < 30.1, taxon_level == "family")

n_southern_families <- southern_families %>% pull(higher_taxon) %>% n_distinct %>% english2
```

@@

```{r limits-stats}
# Find the species with the widest latitudinal distribution
taxon_with_widest_lat <- lat_span_summary %>%
  mutate(lat_span = max_lat - min_lat) %>%
  slice_max(lat_span) %>%
  left_join(green_list, by = "taxon") %>%
  # Parse out species name and author
  mutate(gn_parse_tidy(.$scientific_name) %>% select(canonicalsimple, authorship)) %>%
  mutate(text = glue("*{canonicalsimple}* {authorship}"))

# Find number of species occurring across areas defined by straits
n_widespread_taxa <-
  lat_span_summary %>%
  # Filter by approx location of Tokara strait (29) and Tsugaru Strait (41.5)
  filter(min_lat < 29, max_lat > 41.5) %>%
  pull(taxon) %>%
  n_distinct %>%
  english2
